Turbulence in compact to giant HII regions

### 2022-01-19 strucfunc_paper.tex
Discussion of relation of CO and Ha kinematics

\newcommand\ha{\ensuremath{\text{H}\alpha}}
\newcommand\Wav[1]{\ensuremath{\lambda #1}}

@@

Figure~\ref{fig:velocity-maps} shows maps of the \ha{} centroid velocity
\(V_c\) for each \hii{} region in our sample,
calculated as described above in section~\ref{sec:spectr-datas}
and after subtracting the mean systemic velocity in each case.
The color scheme ranges from the most negative velocities in blue
to the most positive velocities in red,
but the amplitude of the velocity fluctuations differs considerably between regions,
as indicated by the color scale bar on each map,
ranging from \SI{\pm 7}{km.s^{-1}} for the Lagoon
up to  \SI{\pm 40}{km.s^{-1}} for 30~Doradus.
This can be appreciated quantitatively in Figure~\ref{fig:hist},
which shows the probability density function (PDF) of \(V_c\) for each region
on a common velocity scale and arranged in order of decreasing amplitude
of velocity fluctuations. 
All the distributions are far from being consider Gaussian, implying that there are multiple kinematics components in the regions.

@@

>>>>

\newcommand\ha{\ensuremath{\text{H}\alpha}}
\newcommand\Wav[1]{\ensuremath{\lambda #1}}
% Chemical formulae
\newcommand*\chem[1]{\ensuremath{\mathrm{#1}}}
\newcommand\csound{\ensuremath{c_{\text{s}}}}

@@

Figure~\ref{fig:velocity-maps} shows maps of the \ha{} centroid velocity
\(V_c\) for each \hii{} region in our sample,
calculated as described above in section~\ref{sec:spectr-datas}
and after subtracting the mean systemic velocity in each case.
The color scheme ranges from the most negative velocities in blue
to the most positive velocities in red,
but the amplitude of the velocity fluctuations differs considerably between regions,
as indicated by the color scale bar on each map,
ranging from \SI{\pm 7}{km.s^{-1}} for the Lagoon
up to  \SI{\pm 40}{km.s^{-1}} for 30~Doradus.
This is illustrated quantitatively in Figure~\ref{fig:hist},
where the probability density function (PDF) of \(V_c\) for each region
is shown on a common velocity scale and arranged in order of decreasing amplitude
of velocity fluctuations. 
All the distributions are far from being consider Gaussian, implying that there are multiple kinematics components in the regions.

@@

\subsubsection{Kinematic relationship between ionized gas, molecular gas, and stars}
\label{sec:kinem-rela-betw}

Naively, the relatively small observed CO velocity dispersion
(\(\sigma(\chem{CO}) \ll \sigma(\ha)\) in all regions
for both \(\sigma\los\) and \(\sigma\pos\))
might be thought to rule out the molecular gas kinematics
as a \emph{cause} of the ionized gas kinematics.
However, this is not necessarily the case, so long as \emph{some} molecular gas
exists at the higher velocities typical of the ionized gas.

This is because a relatively small column of material
(\(N(\chem{H}) \sim \SI{e20}{cm^{-2}}\))
is sufficient to trap the ionization front,
in which case the \ha{} surface brightness is limited by the
incident ionizing flux
\citetext{the Ferland mechanism,
  see section~5.1 of \citealt{Baldwin:1991a}
  and section~B.2.1 of \citealt{Ferland:2012a}
}.
On the other hand, the bulk of the CO emission will come from much larger
column densities (\(N(\chem{H}) \sim \SI{e22}{cm^{-2}}\)),
where the molecules are effectively shielded from dissociation.
% If a given region can be considered as an ensemble of clouds of
% different masses, then ... would require lower mass clouds to move
% faster though, so needs dynamical relaxation

\textbf{Check  if this is actually feasible before developing further}

Very schematically, one could represent these processes as:
\begin{equation}
  \label{eq:1}
  \sigma^2(\ha) \sim \zeta \sigma^2(\chem{CO}) + \eta
\end{equation}
where \(\zeta\) is a multiplicative factor
due primarily to the saturation at the Ferland limit,
and \(\eta \sim \csound^2\) is an additive factor due to additional turbulence
generated by the transonic ionized photoevaporation flows.

\textbf{Estimate these two factors by fitting to the observations}
